---
title: "Lab6VarmKorv"
author: "Albin Västerlund and Eric Herwin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package contains the three different methods for comuting the knapsack problem. The methods with functions are:

- `brute_force_knapsack`: Performs a brute force computation and you can have the option of parallel-compute this aswell.
- `dynamic_knapsack`: Performes a dynamic computation 
- `greedy_knapsack`: Performes a dreedy method of the knapsack problem

All of these methods are producing a output of a list with the `maximum` value of the opimal weight and `elements` that produces the maximum value. 

More reading about the knapsack problem can be found [here](https://en.wikipedia.org/wiki/Knapsack_problem)


## Question for 1.1.2

How long time does it take to run the brute force search on $n = 16$?

```{r,echo=FALSE,message=FALSE}
library(Lab6VarmKorv)

```
```{r,echo=FALSE,include=FALSE}
set.seed(42)
knapsack_objects <-
data.frame(
  w=sample(1:4000, size = 2000, replace = TRUE),
  v=runif(n = 2000, 0, 10000))
#Brute force


x <- Sys.time()
brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500,parallel = FALSE)
y <- Sys.time()
time <- y-x
```

```{r}
#brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500,parallel = FALSE)
time
```


We can see in the chunk how long time it takes for the brute force function.


## Question for 1.1.3

How much time does it takes to run the algorithm, with the dynamic method, to calculate $n = 500$?


```{r,echo=FALSE,include=FALSE}
x <- Sys.time()
knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500)
y <- Sys.time()
time <- y-x
```

```{r}
#knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500)
time
```



## Question for 1.1.4


How much time does it takes to run the algorithm, with the greedy method, to calculate $n = 1000000$?


```{r,echo=FALSE,include=FALSE}
knapsack_objects <-
data.frame(
  w=sample(1:4000, size = 1000000, replace = TRUE),
  v=runif(n = 2000, 0, 10000))

x <- Sys.time()
lord <- lapply(seq(1001, 1000000, 1000) , function(z){
  greedy_knapsack(x = knapsack_objects[(z-1000)  :z,], W = 3500)
} )
y <- Sys.time()
time <- y-x
```

```{r}
#knapsack_dynamic(x = knapsack_objects[1:1000000,], W = 3500)
time
```


## Question for 1.1.6

So we will use the package `lineprof` to see if the code can be more efficient. To install the package, you have to use the code: `devtools::install_github("hadley/lineprof")`.

### brute force, no parallel

```{r}
#lineprof(hej <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500,parallel = FALSE))
```


In this approach of the brute force method the lineprof funciton found that the `lapply` functions are the ones that slows it down. For this approah, there is probably very little to do to make this approach faster. Mabye to make calls to C++, Python or something equivalent.  

### brute force, with parallel

```{r}
#lineprof(hej <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500,parallel = TRUE))
```

In this function, according to `lineprof` the step that takes the longest, and significant, amount of time is a cluster function `makeCluster` which is needed to set how many cores that are avaliable. Though i do not think you can make this any faster then to just swqitch from the OS Windows to, for example, Linux.

### knapsack_dynamic

```{r}
#lineprof(hej <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500))
```

In this function the time that everything takes to run is very spread out, so basically just to improve the whole `for` and replace with, for example implementing function from the `apply()` family. Though with that approach will be complicated, due to the limitations of indexing with an `apply`.



### greedy_knapsack

```{r}
#lineprof(hej <- greedy_knapsack(x = knapsack_objects[1:800,], W = 3500))
```


In this function it is the sorting that take the most of the time. To make this faster an implementation of, for example the package `dplyr` could be used which is written in the language: C. For the sorting problem you could use the function `arrange()` in the package `dplyr`. 







